---
title: Factor analysis and Linear Regression - the importance of the joint applicability of two data analysis techniques.
author: 
  - name: Fernando Henrique Casalunga
    email: fernandocasalunga@gmail.com
    affiliation: Federal University of Pernambuco
    footnote: CASALUNGA, F. H.
address:
  - code: Federal University of Pernambuco
    address: Department of Political Science (UFPE)
    

journal: ""
date: "`r Sys.Date()`"
output: rticles::elsevier_article

references:
- id: Figueiredo1
  title: What to do and what do not do with regression - assumptions and applications of the linear least squares (OLS) models.
  author: 
  - family: FIGUEIREDO
    given: Dalson
  container - title: Politica Hoje
  Volume: 20
  issue: 1
  publisher:
  type: article-journal
  issued:
    year: 2011

- id: Figueiredo2
  title: Guaranteed Factor Analysis or Your Money Back - An Introduction to Data Reduction.
  author: 
  - family: FIGUEIREDO
    given: Dalson
  container - title: Revista Eletrônica de Ciência Política
  Volume: 5
  issue: 2
  publisher:
  type: article-journal
  issued:
    year: 2014


abstract: |
  The article uses a quantitative exploratory methodology to analyze the 'Quality of Government' database. We used Basic Dataset (2016) to present the application of four multivariate models that combine two techniques of data analysis, factorial and regression, in order to verify the validity of the causal inference of these models against the assumptions indicated in the literature. At the same time, it shows how the combination of these two techniques helps the researcher to identify in a robust way the independent causal effects that affect the phenomenon of researcher interest, in our case the civil liberty in Contemporary States.


---

Introduction
============

```{r library, echo=FALSE, message=FALSE, warning=FALSE} 

# Access to library
if(require(stargazer) == F) install.packages('stargazer'); require(stargazer)
if(require(readr) == F) install.packages('readr'); require(readr)
if(require(ggplot2) == F) install.packages('ggplot2'); require(ggplot2)
if(require(car) == F) install.packages('car'); require(car)
if(require(MASS) == F) install.packages('MASS'); require(MASS)
if(require(xtable) == F) install.packages('xtable'); require(xtable)
if(require(lmtest) == F) install.packages('lmtest'); require(lmtest)
if(require(tidyverse) == F) install.packages('tidyverse'); require(tidyverse)
if(require(foreign) == F) install.packages('foreign'); require(foreign)
if(require(FactoMineR) == F) install.packages('FactoMiner'); require(FactoMineR)
if(require(stats) == F) install.packages('stats'); require(stats)
if(require(reshape2) == F) install.packages('reshape2'); require(reshape2)
if(require(zoo) == F) install.packages('zoo'); require(zoo)
if(require(knitr) == F) install.packages('knitr'); require(knitr)
if(require(xtable) == F) install.packages('xtable'); require(xtable)

# Importing data bse into .csv format - correlation matrix
library(readr)
analysis_data_c <- read_delim("analysis_data_c.csv", 
                            ";", escape_double = FALSE, trim_ws = TRUE)

# Importing database into .csv format - template 1
library(readr)
analysis_data <- read_delim("analysis_data.csv", 
                            ";", escape_double = FALSE, trim_ws = TRUE)

# Importing database into .csv format - template 2
library(readr)
analysis_data_2 <- read_delim("analysis_data_2.csv", 
                              ";", escape_double = FALSE, trim_ws = TRUE)

# Importing database into .csv format - template 3
library(readr)
analysis_data_3 <- read_delim("analysis_data_3.csv", 
                              ";", escape_double = FALSE, trim_ws = TRUE)

# Importing database into .csv format - template 4
library(readr)
analysis_data_4 <- read_delim("analysis_data_4.csv", 
                              ";", escape_double = FALSE, trim_ws = TRUE)

```


```{r base, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Rename correlation matrix
colnames(analysis_data_c) <- c("Civil Liberty 1", "Confidence in Government",
                             "Favoritism in decisions of government officials",
                             "Accountability", "Control of Corruption", 
                             "Ethical behavior of firms", "Property rights", 
                             "Political Stability", "Functioning of government",
                             "Cultural Diversity", 
                             "Ethnic Fractionalization", "Language Fractionalization")

#===========
# Rename database variables model 1
colnames(analysis_data) <- c("Civil Liberty 1", "Confidence in Government",
                             "Favoritism in decisions of government officials",
                             "Accountability", "Control of Corruption", 
                             "Ethical behavior of firms", "Property rights", 
                             "Political Stability", "Functioning of government",
                             "Cultural Diversity", 
                             "Ethnic Fractionalization", "Language Fractionalization", 
                             "Institucional Efficiency", "Ethnic_linguistic Plurality")

#============
# Rename database variables model 2 
colnames(analysis_data_2) <- c("Civil Liberty 1",
                             "Favoritism in decisions of government officials",
                             "Accountability", "Control of Corruption", 
                             "Ethical behavior of firms", "Property rights", 
                             "Political Stability", "Functioning of government",
                             "Cultural Diversity", 
                             "Ethnic Fractionalization", "Language Fractionalization", 
                             "Institucional Efficiency", "Ethnic_linguistic Plurality")


#============
# Rename database variables model 3 
colnames(analysis_data_3) <- c("Civil Liberty 2", "Confidence in Government",
                             "Favoritism in decisions of government officials",
                             "Accountability", "Control of Corruption", 
                             "Ethical behavior of firms", "Property rights", 
                             "Political Stability", "Functioning of government",
                             "Cultural Diversity", 
                             "Ethnic Fractionalization", "Language Fractionalization", 
                             "Institucional Efficiency", "Ethnic_linguistic Plurality")

#============
# Rename database variables model 4
colnames(analysis_data_4) <- c("Civil Liberty 2",
                             "Favoritism in decisions of government officials",
                             "Accountability", "Control of Corruption", 
                             "Ethical behavior of firms", "Property rights", 
                             "Political Stability", "Functioning of government",
                             "Cultural Diversity", 
                             "Ethnic Fractionalization", "Language Fractionalization", 
                             "Institucional Efficiency", "Ethnic_linguistic Plurality")

```

```{r savedataset, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#=============================
# Save correlation matrix database
write.table(analysis_data_c, "analysis_data_c.csv", sep = ";", row.names = F)

# Save database models 1 through 4 with renamed variables
# model 1
write.table(analysis_data, "analysis_data.csv", sep = ";", row.names = F)

# model 2
write.table(analysis_data_2, "analysis_data_2.csv", sep = ";", row.names = F)

# model 3
write.table(analysis_data_3, "analysis_data_3.csv", sep = ";", row.names = F)

# model 4
write.table(analysis_data_4, "analysis_data_4.csv", sep = ";", row.names = F)

```

The article explores the data from the "Quality of Government Basic Dataset" (2016) [^1] based on our research design, we used two techniques of quantitative data analysis, factorial analysis and the regression of ordinary least squares (OLS) ) [^2] to measure the effect of some independent variables that we believe can explain how the phenomenon of civil liberty in contemporary states behaves.

[^1]: The Quality of Government Basic Dataset, version Jan16. University of Gothenburg: The Quality of Government Institute. Available in http: // www.qog.pol.gu.sedoi:10.18157/QoGBasJan16>. Acess in 20.12.2018.

[^2]: For a deeper understanding of the assumptions of the techniques used see also: "Factorial Analysis: An Introduction to Data Reduction" rev. eletrônica de Ciência Política V. 5, nº2 (2014); "What to do with regression: assumptions and applications of the linear model (OLS)" rev. eletrônica Política Hoje, vol. 20, n.1 (2011).

Regarding sample size, during the planning of the measurement of variables, four models were chosen, two of them with larger samples (N = 127) (1 and 3) and another two (2 and 4) with smaller samples (N = 51). For the purpose of comparison, we analyzed the results identified in each of these models according to ten assumptions of the regression models indicated by @Figueiredo1,  to present the one that best fits the explanation of our phenomena of interest, civil liberty.

Thus, the unit of analysis of this study is the countries, models 1 and 3 have information for 51 of them, while models 2 and 4 have 127 observations. These models differ according to the variation in omitted cases, which occurs due to the absence of data for one of our tested independent variables 'government confidence' which is measured by opinion polls conducted by World Values Surveys in 76 countries.

Furthermore, models 1 and 2 differ from models 3 and 4 as regards the dependent variable used to measure civil liberty. In models 1 and 2 we measured the effects of the independent variables on civil liberty taken as a categorical dependent variable with values varying from 1 to 7 -where 1 represents the countries with the highest and 7 those with the lowest civil liberty-, already with models 3 and 4 we measured the effects of the independent variables on civil liberty as a continuous dependent variable -with values varying from 1 to 10, with 1 being the value assigned to countries with low and 10 with high levels of civil liberty-.

Thus, the article is divided into three sections, the first presents step by step the construction of the democratic inclusiveness index (IDI), composed of two latent variables [^3] named institutional efficiency and ethnic-linguistic plurality which were transformed using the technique of factorial analysis; the second section uses the two factorial indicators that make up our construct as independent variables in four models of multivariate linear regression, to which we add three more variables that we believe exert an intervening effect on our variable of interest, civil liberty; the third section summarizes the results of the four models and the fourth section presents a brief concluding discussion.

[^3]: Definition of latent variable: corresponds to a variable that contains several variables that theoretically make up the created factor, the way in which it is possible to measure a certain phenomenon of interest @Figueiredo2.

Regarding the transparency and replicability of this study, all statistical analyzes were performed with the aid of RStudio software [^4] to analyze statistical data and RMarkdown [^5] for documentation, computational routines were duly reported through the organization of the TIER protocol [^6] and versioning was done with Github platform [^7], the data and codes for replication of the results were made available in the Dataverse repository of Harvard University. We also chose to prioritize the presentation of the viability of the constructs and the results of the models through a graphical interface, because we believe that in this way we facilitate the reader's understanding of our results, all the tables adjacent to the results are attached.

[^4]: RSTUDIO, version 1.1.463. Available in <https://www.rstudio.com/products/>. Acess in 15.12.2018.
[^5]: RMarkdown, version 1.11. Availabe in <https://rmarkdown.rstudio.com/>. Acess in 15.12.2018.
[^6]: The TIER Documentation Protocol provides instructions for assembling a set of electronic files that document all the steps of data processing and analysis you conduct for an empirical research paper. Disponível em <https://www.haverford.edu/project-tier/protocol-v2>. Acess in 16.07.2018
[^7]: GitHub. Available in <https://github.com/casalungafh/artigo_ad_casalunga_fernando>. Acess in 10.01.2019.

Sections
===============

## Section 1: Transformation of independent parameters into latent variables.

The analysis of this study follows an exploratory pattern of data, a technique recommended for research in the initial stages, according to this perspective, the following multicollinear variables were identified: X1) control of corruption; X2) ethical behavior of firms; X3) property rights; X4) political stability; X5) government functioning; X6) cultural diversity; X7) ethnic fractionalization; and X8) linguistic fractionalization. Due to the high degree of correlation between these variables, it is advisable to transform its main components into latent variables, composed of the main information present in each of them in a synthetic way, the technique used to carry out the transformation is called factorial analysis [^8], and will be developed below.

[^8]: Factor analysis as a data reduction technique is used to construct indexes that synthesize information from a large number of variables into a smaller number of factors / components. The correlation pattern observed among the original variables is condensed into a single variable that contains the factors in common between them @Figueiredo1.

Thus, factor construction depends on the level of correlation between the selected independent variables. This is because the factorial analysis uses the common variance between the correlated variables ($X1, X2, X3 ... X8$) to construct a factorial indicator that contain information of all the observed variables divided into latent factors, so with this technique we estimate linear combinations of the original variables from the variance in the correlation pattern between the variables.

In relation to the correlation between the independent variables that compose the factors that will be constructed, the literature recommends that most of the coefficients be higher than ($R \ge 0.300$), independent of the signal, in this way, "the higher the level of correlation between the the most appropriate variables will be the different techniques of data reduction "(TABACHNICK and FIDELL, (2007), apud [@Figueiredo2: p. 190]. [^9] 

Graph 1 represents the high correlation pattern among the variables, indicating that the matrix of data is adequate to the use of factorial analysis.

[^9]:In the planning of factorial analysis, the literature indicates that the following stages are observed: a) definition of the theoretical dimension of the latent variable; b) identification of observed variables caused by the latent variable, c) data collection; d) computational transformation; e) discussion of the results [@Figueiredo2: p. 187].

```{r analyze1, echo=FALSE, message=FALSE, warning=FALSE, results='asis'} 

# Generate correlation matrix - model 1
cormat <- cor(analysis_data_c[, -c(1:4)])
melted_cormat <- melt(cormat)

# Correlation matrix graph - model 1
ggplot(data = melted_cormat, 
       aes(x=`Var1`, y=`Var2`, fill=value)) +
  geom_tile()

```

In the annex, the descriptive statistics of the independent variables ($X1, X2 ... X8$) used to compose the factors are presented in detail, tables 1 and 2 present the indicators that were used in the construction of the factorial indicators 'Institutional Efficiency' (factor 1) and 'Ethnic-linguistic Plurality' (factor 2).

Tables 3 and 4 present, respectively, the adequacy statistics of the sample and the communalities. At Table 3 we have the adjusted Kaiser-Meyer-Olkin (KMO) statistic that indicates the proportion of the variance attributed to the factors that will be transformed, the indicator varies from 0 to 1, the higher the assumed value, the more accurate is the result of the analysis factorial. In our analysis the ($KMO = 0.860$) indicates that we can proceed with the application of the technique.

Graph 2 shows the total variance of each factor that will be transformed, the variability of the variables and the accumulated percentage of each variable extracted for the construction of the factors, while graph 3, contains the information in a scree plot. Both objectively represent how many factors are to be extracted, the dotted line signals the cut level of the Kaiser rule, the abrupt drop signals for transformation of the independent variables into two constructs.

```{r analyze2, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

# Factorial Analysis (PCA)
#=======================================
# Model 1
pca1 <- princomp(analysis_data[, -c(1:4)], scores=TRUE, cor=TRUE)
# Sumario do PCA
summary(pca1)

# Loading main Components
loadings(pca1)

par(mfrow=c(1,2))
# Checking how many factors should we use
plot(pca1) # cotovelo em 2 fatores

# Selecting the 2 factors - template 1
analysis_data$fac1 <- pca1$scores[,1]
analysis_data$fac2 <- pca1$scores[,2]

# Scree plot of eigenvalues
screeplot(pca1, type="line", main="Scree Plot")
abline(1,0, lty=2)

```

From the way the extracted components and observed variables relate, the next step will be to produce a rotated component matrix. The rotation seeks to facilitate the interpretation of the observed solution without changing its original mathematical properties (TABACHNICK and FIDELL, (2007), apud, [@Figueiredo2: p. 191], the orthogonal rotation assumes that the correlation between the dimensions is zero, indicating that they are statistically independent (orthogonal).

```{r analyze3, eval=FALSE, echo=FALSE}

# Rotation varimax (pca1 $ rotation)
# Factor analysis 
fatorial_varimax <- factanal(analysis_data[, -c(1:4,13:14)],factors=2, rotation="varimax", scores="regression")

```

When applying the technique of factorial analysis we determine the type of rotation (orthogonal varimax), and the amount of factors that must be extracted, in this case, two. Thus, the two factors that will be constructed will be composed of two components, the first representing 51% and the second 28% of the total variance of the independent variables. Together the two accumulated values add up to 80% of the total information of the independent variables. 

In the annex, table 5 summarizes the results of the accumulated variance, table 6 present the rotational component matrix of the two components, the uniqueness of the variables, the degree of freedom and the chi-square test of the factorial analysis and table 7 shows the average variation of the factors created, 'Institutional Efficiency' and 'Plurality Ethnic-linguistic' databases containing information for 127 countries in the QOG database.

With the ownership of the transformed factors, they become a construct, or indicator that we call Inclusive Democracy Index (IDI). To better identify the operationalization of these factors in our construct, we represent through the graphical interface the dispersion of all the countries of our database according to the two latent variables transformed.


```{r analyze4, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

# Descriptive of factors 1 and 2
summary(analysis_data$`Institucional Efficiency`)
summary(analysis_data$`Ethnic_linguistic Plurality`)

ggplot(analysis_data, aes(x = `Ethnic_linguistic Plurality`, y = `Institucional Efficiency`)) +
  geom_point() +
  geom_vline(xintercept = 0, size =1, linetype = "dashed") +
  geom_hline(yintercept = 0, size =1, linetype = "dashed") +
  labs(x = "Ethnic_linguistic Plurality", y = "Institutional Efficiency")

```


```{r analyze5, eval=FALSE, echo=FALSE}

ggplot(analysis_data, aes(x = Institucional Efficiency, y = Ethnic-linguistic Plurality)) +
  geom_point()+
  geom_text(aes(label = analysis_data$NOMEDOPAIS)) +
geom_smooth(method='lm')

```

In this way, we can identify the analysis units by quadrant. In the quadrant are countries with low institutional efficiency and high ethnic-linguistic plurality; already quadrant 2 presents countries with high ethnic-linguistic plurality and high institutional efficiency; in quadrant 3 are countries with low institutional efficiency and low ethnic-linguistic plurality; and the quadrant 4 countries with high institutional efficiency and low ethnic-linguistic plurality.

In the next section, we will use the two factors transformed in this section as independent variables in four multivariate regression models, in order to measure the effects produced on our variable of interest, civil liberty, in the countries of the QOG database (2016).[^10]

[^10]: The Quality of Government Institute. QoG Basic Cross-Section Datase. Availble in <https://qog.pol.gu.se/data/datadownloads/data-archive>. Acess in 05.12.2018.


## Section 2 - Construction of Multivariate Linear Regression (MQO)


The basic structure of a linear regression model is given by the average variation of a dependent variable ($Y$) according to the average variation of the other independent variables. The literature shows that in order to use the ordinary least squares (OLS) model, the researcher must comply with a series of recommendations that help increase the validity and reliability of the results, attention to the assumptions of the models is crucial, since the analysis of regression became the statistical model most commonly employed in contemporary political science [@Figueiredo1: p. 46].

The ten recommendations of the MQO models follow in the wake of the Lewis-Beck (1980) and Kennedy (2009) postulates, the first one says that the relationship between the dependent variable and the independent variables must be linear, the second prevents against measurement errors of the variables, the third proposes that the mean of the error term should be equal to zero, the fourth that the error term be constant according to the distribution of the values of the independent variable, the fifth that there should be no autocorrelation between the error terms, the sixth that the independent variable should not be correlated with the error term, the seventh that no relevant variable to explain ($Y$) was excluded from the model, the eighth that there is no high correlation between the independent variables, the ninth that the distribution of the error is normal, and the tenth one has an adequate proportion between number of cases and estimated parameters [@Figueiredo1: p. 52].

Regarding the application of the linear regression model, the eighth indication is salutary in indicating that we should not use highly correlated variables, so called multicollinearity occurs when high levels of correlation are observed between more than two independent variables, their main effect is to affect the efficiency of the estimates, in this case, we tend to overestimate the magnitude of the standard errors of the regression coefficients, which end up impairing the reliability of the significance tests (p-value) and confidence intervals [@Figueiredo1].

Regarding the purpose of this study, to demonstrate the importance of the joint use between the techniques of factorial analysis and linear regression, according to the indications, it is evident the importance of techniques to reduce the dimensionality of the data, because with the use of the analysis factorial the researcher is able to transform highly correlated independent variables into latent constructs that carry the condensed information of the original variables, which helps him to construct more robust models that contain a great amount of information but at the same time circumvent the problem of multicollinearity. For this reason, we emphasize the importance of combining the techniques of factorial analysis with linear regression, the disadvantage of this procedure is the impossibility of observing the individual effect of each explanatory variable of the model [@Figueiredo1: p 56].

It should also be noted that, in addition to the problems related to the multicollinear independent variables, the use of small samples can also affect the results of the linear regression models, to test this assumption we used models with 51 and 127 cases.

Thus, it is recommended that the researcher elaborate models capable of minimizing the information of the variables and maximizing the number of observations, attention to these procedures helps to increase the degree of reliability and the possibility of generalization of the results.

Specifically, in regression models the dependent variable ($Y$) represents the phenomenon we seek to understand, explain or predict. The constant, ($\alpha$), represents the best estimate of ($Y$) in the absence of independent variables. The independent variables are represented by $X1 ... + X5$, the coefficients ($\beta$) represent the average variation observed in ($Y$) when a certain independent variable is raised in one unit, keeping the others constant. For example, ($X1$) represents the variation observed in ($Y$) by raising ($X1$) in one unit, keeping the remaining variables constant. The model error, ($\epsilon$), represents all that can affect ($Y$), but can not be explained by the model. To fix the interpretation, we chose to present the equation of the mathematical model: 

$Y = \alpha + \beta_1 x + \beta_2 x2 + \beta_3 x3 + \beta_4 x4 + \beta_5 x5 + \epsilon$ 


In this study, models 1 and 2 use a categorical dependent variable. For models 3 and 4 a continuous dependent variable is used for comparison purposes, because the literature indicates that we should opt preferentially for the use of quantitative dependent variables, discrete or continuous [@Figueiredo1: p. 61].

In annex (2) we describe the measurement of the dependent variable and the other independent variables used in the model to explain our phenomenon of interest, civil ciberty. Table 8 shows the dependent variable ($Y$) that the model intends to explain from the independent variables ($X1, X2, .. + X5$). Also, the hypothetical example seeks to understand the impact of the independent variables ($X1, X2 ... + X5$) on the dependent variable ($Y$).

The applied models seek, therefore, to understand the causal effect on the civil Liberty produced by the variation of public confidence in government, favoritism in official government decisions, accountability, institutional efficiency and ethnic-linguistic plurality, the last two are the multicollinear factors that make up a construct denominated index (IDI) produced in section 1 of this article.

In this way the multivariate linear regression models attempt to measure the degree of civil liberty, respectively, in 51 (models 1 and 3) and 127 countries (models 2 and 4) of the QOG Basic database (2016) through four regression models linear multivariate analysis. In the annex, table 9 summarizes the descriptive statistics of the independent variables.
Thus, the hypotheses that will be tested with the multivariate models are listed, our objective is to observe if the effect produced by the independent variables on civil liberty is significant and its magnitude. Therefore, it is stated:

H0: indicates that there is no significant relationship between public confidence in government ($X1$); favoritism in government decisions ($X2$); accountability ($X3$); ($X5$) and civil liberty ($Y$), that is, null hypotheses indicate that the effect is null ($\beta = 0$)

Ha: indicates that there is a significant relationship between ($X1... + X5$) and $Y$, that is, alternative hypotheses indicate that the effect is positive ($\beta \neq 0$).

However, in our models, the interpretation of the hypothesis tests of ($X2$) will be different from the others, we expect the inverse effect for this variable in our dependent variable, that is, the greater the favoritism in government decisions the lower the degree of civil liberty in that country ($\beta <0$).

In the following section, we summarize the results of the four models based on the assumptions of the linear regression models indicated by the literature [@Figueiredo1; @Figueiredo2].

## Section 3 - Results of multivariate linear regression models

Considering the previous section, it is worth mentioning that in our models, the inclusion of the tested variable 'Trust in government' ($X1$) represented by public opinion polls conducted by the World Values Survey (WVS) reduces the number of cases of 127 (models 2 and 4 ) to 51 (models 1 and 3) due to the lack of measurement of this variable in 76 countries for which WVS data are not available. Another important point is the use of two distinct dependent variables to verify the phenomenon of civil liberty, a categorical one ranging from 1 to 7 where the smallest number indicates the countries with the greatest civil liberty, and another continuous one ranging from 1 to 10 where the smallest number indicates the countries with the lowest civil liberties.

Table 10 presents the summary of the four multivariate regression models applied to the study of civil liberty in the countries of the QOG database (2016), we observe the degree of explanation of the models through the $R2-adjusted$. The table summarizes the results of the F statistical tests, the $p-value$, the degrees of freedom and the standard deviation of the models. This information indicates whether or not the estimates are statistically significant, that is, the degree of validity of the inference production depends on the tests of significance. The standard error is a measure of variability of the estimate, that is, the smaller the standard error, the more accurate the estimate is, so if the standard error is too large, the estimate will not be significant.

From the analysis of the models it was identified that the model 1 has the greatest explanatory power representing that the chosen independent variables explain 85% of the 51 cases ($R2-adjusted = 0.85$) when we use civil liberty as a categorical dependent variable, already in model 3 in which we employ civil liberty as a continuous variable the explanatory power for 51 cases decreases to 71% ($R2-adjusted = 0.711$). We also observed that models 2 and 4 follow a similar pattern, and in both cases we observed 127 cases, (we include the countries for which there was no information on WVS public opinion surveys), model 2 that uses the categorical dependent variable can explain 83% of the 127 cases ($R2-adjusted = 0.832$), while model 4 with the continuous dependent variable explains 76% of the cases ($R2-adjusted = 0.763$).

As mentioned above, in order to interpret the results of model 1 and 2 it is necessary to take into account that the indicator of the categorical variable indicates that countries with greater civil liberty receive the value 1 while those with less freedom receive a value equal to 7.

Thus, model 1 results indicate that, on average, by raising the level of public confidence in government in one unit, it is expected on average a reduction of 1,002 points in civil liberties ($p-value < .01$). Similarly, on average, favoritism increases in official government decisions, on average, a reduction of 0.561 points in civil liberty ($p-value <.05$), the occurs with the variables $X3$, $X4$ and $X5$. Unless we see the increase in unity in accountability has a positive effect on civil liberty, it is expected that the variation will increase by an average of 0.052 civil liberties ($p-value <.01$), institutional efficiency has a similar effect, it is expected that the variation of one unit in this factor causes, on average, an increase of 0.496 in civil liberty ($p-value <.01$), while ethnic-religious plurality has a positive effect, however not statistically significant on civil liberties in this model.

The findings of model 2, with 127 cases, indicate that, on average, favoritism increases in official government decisions should result in a reduction of 0.675 points in civil liberty ($p-value <.01$), the inverse occurs with variables $X3$, $X4$ and $X5$, as we can see the increase of one unit in the transparency of public accounts (accountability) has a positive effect on civil liberty, it is expected that the variation causes, on average, an increase of 0.067 in freedoms ($p-value <.01$), institutional efficiency has again a positive effect, it is expected that the variation of one unit in this factor causes, on average, an increase of 0.397 in civil liberty ($p-value <.01$), othwerwise ethnic-religious plurality has a positive but not statistically significant effect on civil liberties in this model.

Considering that, in order to interpret the results of model 3 and 4, it must be taken into account that the indicator of the dependent variable is now continuous, so that countries with greater civil liberty receive the value 10, while those with less freedom receive an equal value 1.

The results of model 3, with 51 cases, indicate that, on average, by raising the level of confidence of the public in government in one unit is expected a reduction of 1,380 points in the civil liberty ($p-value <.05$), similarly, on average, when favoritism in official government decisions is raised is expected a reduction of 0.607 in the dependent variable, but this result is not statistically significant. Again, the reverse is true of variables $X3$, $X4$, and $X5$, so increasing a unit in the transparency of accountability has a positive effect on civil liberty, it is expected that the variation causes, on average, a increase of 0.074 in civil liberties ($p-value <.01$), we can also see that institutional efficiency has a similar effect, it is expected that the variation of one unit in this factor causes, on average, an increase of 0.556 in civil liberty ($p-value <.05$), while again ethnic-religious plurality has a positive but not statistically significant effect on civil liberties in this model.

The findings of model 4, with 127 cases, indicate that, on average, favoritism increases in official government decisions produces a reduction of 0.921 points in civil liberty ($p-value <.01$), again, the reverse occurs with variables $X3$, $X4$ and $X5$, and the increase of one unit in the transparency of public accounts (accountability) has a positive effect on civil liberty, we expected that, on average, the variation of this variable causes an increase of 0.104 in civil liberties ($p-value <.01$), institutional efficiency has a similar effect, we expected that, on average, the variation of one unit in this factor causes an increase of 0.486 in civil liberty ($p-value <.01$), ethnic-religious plurality has a positive but not statistically significant effect on civil liberties in this model.

On the one hand, the results pointed out that in the four models the first latent transformed variable works as expected, thus we can say that institutional efficiency has a causal effect on the civil liberties of the countries, however on the other the second latent trasnformed variable, ethnic-linguistic plurality does not seem to be significant to explain the increase of civil liberties in the countries. Graphs 5, 6, 7 and 8 summarize the confidence intervals and indicate the statistical significance of the results and the standard erros range clearly.

```{r analyze6, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Multivariate Linear Regression Model 1

model_1 <- lm(`Civil Liberty 1` ~ `Confidence in Government` +
                 `Favoritism in decisions of government officials` + 
                 `Accountability` + 
                 `Institucional Efficiency` + 
                 `Ethnic_linguistic Plurality` , 
               data = analysis_data)

#=================
# Multivariate Linear Regression Model 2 - drop Confidence in Government

model_2 <- lm(`Civil Liberty 1` ~  +
                 `Favoritism in decisions of government officials` + 
                 `Accountability` + 
                 `Institucional Efficiency` + 
                 `Ethnic_linguistic Plurality` , 
               data = analysis_data_2)

#============
# Multivariate Linear Regression Model 3
model_3 <- lm(`Civil Liberty 2` ~ `Confidence in Government` +
                 `Favoritism in decisions of government officials` + 
                 `Accountability` + 
                 `Institucional Efficiency` + 
                 `Ethnic_linguistic Plurality` , 
               data = analysis_data_3)

#==============
# Multivariate Linear Regression Model 4 - drop Confidence in Government
model_4 <- lm(`Civil Liberty 2` ~  +
                 `Favoritism in decisions of government officials` + 
                 `Accountability` + 
                 `Institucional Efficiency` + 
                 `Ethnic_linguistic Plurality` , 
               data = analysis_data_4)

# Graph for the representation of the confidence intervals of the estimated coefficients - model 1
par(mfrow=c(1,1))
betas <- coefficients(model_1) 
IC <-  confint(model_1, level=0.95) 

y.axis <- seq(from=1, to=length(betas))
plot(betas, y.axis, type="p", pch=19, xlab="Magnitude of Coefficients", ylab="", 
     axes=F, xlim=c(min(IC-.4), max(IC+.4)), ylim=c(min(y.axis-.2), max(y.axis+.2)), 
     cex=1,yaxs="i",xaxs="i")
segments(IC[,1], y.axis, IC[,2], y.axis)
axis(1, at=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), 
     labels=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), tick=T, cex.axis=1, 
     mgp=c(2,.7,0))
axis(2, at=y.axis, label=names(betas), las=1, tick=T, line=-.5, cex.axis=1, 
     mgp=c(2,.7,0))
abline(v=0, lty=2, col="red")


# Graph for the representation of the confidence intervals of the estimated coefficients - model 2
par(mfrow=c(1,1))
betas <- coefficients(model_2) 
IC <-  confint(model_2, level=0.95) 

y.axis <- seq(from=1, to=length(betas))
plot(betas, y.axis, type="p", pch=19, xlab="Magnitude of Coefficients", ylab="", 
     axes=F, xlim=c(min(IC-.4), max(IC+.4)), ylim=c(min(y.axis-.2), max(y.axis+.2)), 
     cex=1,yaxs="i",xaxs="i")
segments(IC[,1], y.axis, IC[,2], y.axis)
axis(1, at=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), 
     labels=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), tick=T, cex.axis=1, 
     mgp=c(2,.7,0))
axis(2, at=y.axis, label=names(betas), las=1, tick=T, line=-.5, cex.axis=1, 
     mgp=c(2,.7,0))
abline(v=0, lty=2, col="red")

# Graph for the representation of the confidence intervals of the estimated coefficients - model 3
par(mfrow=c(1,1))
betas <- coefficients(model_3) 
IC <-  confint(model_3, level=0.95) 

y.axis <- seq(from=1, to=length(betas))
plot(betas, y.axis, type="p", pch=19, xlab="Magnitude of Coefficients", ylab="", 
     axes=F, xlim=c(min(IC-.4), max(IC+.4)), ylim=c(min(y.axis-.2), max(y.axis+.2)), 
     cex=1,yaxs="i",xaxs="i")
segments(IC[,1], y.axis, IC[,2], y.axis)
axis(1, at=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), 
     labels=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), tick=T, cex.axis=1, 
     mgp=c(2,.7,0))
axis(2, at=y.axis, label=names(betas), las=1, tick=T, line=-.5, cex.axis=1, 
     mgp=c(2,.7,0))
abline(v=0, lty=2, col="red")

# Graph for the representation of the confidence intervals of the estimated coefficients - model 3
par(mfrow=c(1,1))
betas <- coefficients(model_4) 
IC <-  confint(model_4, level=0.95) 

y.axis <- seq(from=1, to=length(betas))
plot(betas, y.axis, type="p", pch=19, xlab="Magnitude of Coefficients", ylab="", 
     axes=F, xlim=c(min(IC-.4), max(IC+.4)), ylim=c(min(y.axis-.2), max(y.axis+.2)), 
     cex=1,yaxs="i",xaxs="i")
segments(IC[,1], y.axis, IC[,2], y.axis)
axis(1, at=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), 
     labels=seq(round(min(IC-.9)), round(max(IC+.9)), by=0.1), tick=T, cex.axis=1, 
     mgp=c(2,.7,0))
axis(2, at=y.axis, label=names(betas), las=1, tick=T, line=-.5, cex.axis=1, 
     mgp=c(2,.7,0))
abline(v=0, lty=2, col="red")


```

In light of our findings, at first glance we may be led to assume that model 1 that presents the highest degree of explanation of the cases in our data matrix is the one that best explains the phenomenon of civil liberty. However, in the next section, we discuss the robustness of the four models against the ten assumptions of the regression models to know if we can confirm or not the expectations regarding model 1.


## Section 4 Discussion and Conclusion

Therefore, since we use the secondary database from the Quality of Government Institute (QOG) repository, we assume that the first second and third assumptions of the regression models indicate, respectively, the linear relationship between the independent parameters and the dependent variable, the importance of the absence of measurement errors in the variables and the importance of the mean value of the error term to be equal to zero were satisfied in our four models, which guarantees the reliability and validity of the parameters used, as well as the consistency of the tests of statistical significance applied.

As for the other assumptions, starting with the models using the categorical dependent variable, it was identified that models 1 and 2 satisfy the fourth assumption that the difference between the observed results and the results predicted by the model should vary uniformly [@Figueiredo1: p. 54]. Graphs 9 and 10 show the homoscedasticity of the residual distribution of these models.

```{r analyze7, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

par(mfrow=c(1,2))
### Evaluating homoscedasticity ###
# Residual analysis of multivariate linear regression model 1
analysis_data$Predicted <- predict(model_1)
analysis_data$Residuo <- analysis_data$`Civil Liberty 1` - analysis_data$Predicted

# Test of the mean value of the error term equal to zero
summary(analysis_data$Residuo)

# Scatter plot of standard residues - model 1 
plot(analysis_data$Predicted, analysis_data$Residuo, pch=21, bg="red", col="red")
abline(0,0, lty = 2)

### Evaluating homoscedasticity ###
# Residual analysis of multivariate linear regression model 2
analysis_data_2$Predicted <- predict(model_2)
analysis_data_2$Residuo <- analysis_data_2$`Civil Liberty 1` - analysis_data_2$Predicted

# Test of the mean value of the error term equal to zero
summary(analysis_data_2$Residuo)

# Scatter plot of standard residues - model 2
plot(analysis_data_2$Predicted, analysis_data_2$Residuo, pch=21, bg="red", col="red")
abline(0,0, lty = 2)


```

According to the fifth assumption of the regression models [@Figueiredo1: p. 56] we test the absence of autocorrelation between the cases using the Durbin-Watson test that detects the presence of autocorrelation in the data, the indicator varies between 0 and 4, values between 1.5 and 2.5 indicate independence of the observation. 

```{r analyze8, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
### Test for Collinearity ###
vif(model_1) # variance inflation factors 
sqrt(vif(model_1)) > 3
sqrt(vif(model_1)) < 1

### Test for autocorrelation of errors 
durbinWatsonTest(model_1)

# Correlation between the dependent variable and the factor 1 - independent variable constructed
cor(analysis_data$`Civil Liberty 1`, analysis_data$`Institucional Efficiency`)

# Correlation between the dependent variable and the factor 2 - independent variable constructed
cor(analysis_data$`Civil Liberty 1`, analysis_data$`Ethnic_linguistic Plurality`)

### Test for non-constant error variance
ncvTest(model_1)

### Test for Collinearity 
vif(model_2) # variance inflation factors 
sqrt(vif(model_2)) > 3
sqrt(vif(model_2)) < 1

### Test for autocorrelation of errors 
durbinWatsonTest(model_2)

# Correlation between the dependent variable and the factor 1 - independent variable constructed
cor(analysis_data_2$`Civil Liberty 1`, analysis_data_2$`Institucional Efficiency`)

# Correlation between the dependent variable and the factor 2 - independent variable constructed
cor(analysis_data_2$`Civil Liberty 1`, analysis_data_2$`Ethnic_linguistic Plurality`)

### Teste para variancia do erro nao constante
ncvTest(model_2)

```

Model 1 presents ($DWT = 1.8$), and model 2 ($DWT = 1.9$) also tested the correlation between latent variables like institutional efficiency (model 1 $R = -0.713$; model 2 $R = -0.697$) and Ethnic-linguistic Plurality (model 1 $R = -0.113$; model 2 $R =-0.242$) with the dependent variable civil liberty, the results indicate that the observations are linear and independent and, therefore, there is no correlation between the error terms, thus guaranteeing the reliability of the tests of significance and confidence intervals. We also tested the variance of the factors (model 1 $chi-square = 2.7$ with 1 degree of freedom; model 2 $chi-square = 8.14$ with 1 degree of freedom).

The sixth assumption of the regression models is difficult to match, the literature indicates that the correlation between the independent variables and the error term may bias the results, however, since this is not an experimental study, it is not possible to manipulate the value of the independent variable, so we chose to violate this assumption, a common measure adopted by quasi-experimental designs such as multivariate linear models. Graphs 11 and 12 indicate that the distribution of the error term for each individual-mode component follows a heteroscedastic pattern, that is, there is a correlation between the error term and the independent variables, with the exception of the latent ethnic-linguistic plurality variable that follows next to a homocedastic distribution but the effect of this variable is not statistically significant.

```{r analyze 9, eval=FALSE, echo=FALSE}

## Non-linearity test
# component added to waste 
crPlots(model_1)

## Non-linearity test
# component added to waste 
crPlots(model_2)

```

The seventh and eighth assumptions of the regression models that point to possible problems in the estimates due to the inadequate specification of the model, such as the inclusion of theoretically irrelevant independent variables that reduce the efficiency of the estimators, or the multicollinearity between the independent variables and the variable depending on the model that may lead to increased variance in the magnitude of the estimated parameters, were satisfied by models 1 and 2, while they include two latent variables as functional mechanisms to avoid the bias imposed by possible omitted variables and, at the same time, to condense the information of the highly correlated variables in a single factorial indicator.

The ninth assumption, can be observed from the analysis of the distribution of the residuals of the models, the amount of information about the phenomenon of civil liberty that can not be explained by our research design should follow a symmetrical distribution to satisfy this assumption. Both models do not satisfy the ninth assumption, graphs 13 and 14 demonstrate that the values of standardized residues of models 1 and 2 do not distribute in a perfectly symmetrical way.

```{r analyze10, echo=FALSE, message=FALSE, warning=FALSE}

### Evaluating waste distribution ###
# Scatter histogram of two residues model 1
par(mfrow=c(1,2))
hist(analysis_data$Residuo, freq=FALSE, ylim=c(0,0.6), main="Distribution of standard errors model 1")
xfit <- seq(min(analysis_data$Residuo),max(analysis_data$Residuo),length=40) 
yfit <- dnorm(xfit) 
lines(xfit, yfit)

# Scatter histogram of two residues model 2
hist(analysis_data_2$Residuo, freq=FALSE, ylim=c(0,0.6), main="Distribution of standard errors model 1")
xfit <- seq(min(analysis_data_2$Residuo),max(analysis_data_2$Residuo),length=40) 
yfit <- dnorm(xfit) 
lines(xfit, yfit)

```

The tenth and final assumption regarding the Central Limit Theorem that proposes to increase the number of cases against the estimated parameters was tested by increasing the number of cases from 51 (model 1) to 127 (model 2), removing the variable confidence in government. In the annex, the result observed in table 10 indicates that the explanatory power of model 2 is smaller when compared to the previous one in terms of $R-adjusted$ and degrees of freedom, therefore, in this case the increase in the number of cases had little influence on the results of the model 2.

In relation to models 3 and 4 that use the continuous dependent variable to measure civil liberty, the analysis indicates homocedastic distribution of the residues against the predicted values. Graphs 15 and 16 summarize the uniform dispersion of the difference between the observed results and the results predicted by the models, therefore both satisfy the fourth assumption.

```{r analyze11, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

par(mfrow=c(1,2))
### Evaluating homoscedasticity ###
# Residual analysis of the multivariate linear regression model 3
analysis_data_3$Predicted <- predict(model_3)
analysis_data_3$Residuo <- analysis_data_3$`Civil Liberty 2` - analysis_data_3$Predicted

# Test of the mean value of the error term equal to zero
summary(analysis_data_3$Residuo)

# Scatter plot of two residues model 3
plot(analysis_data_3$Predicted, analysis_data_3$Residuo, pch=21, bg="red", col="red")
abline(0,0, lty = 2)

### Evaluating homoscedasticity ###
# Residual analysis of the multivariate linear regression model 4
analysis_data_4$Predicted <- predict(model_4)
analysis_data_4$Residuo <- analysis_data_4$`Civil Liberty 2` - analysis_data_4$Predicted

# Test of the mean value of the error term equal to zero
summary(analysis_data_4$Residuo)

# Scatter plot of two residues model 4
plot(analysis_data_4$Predicted
     , analysis_data_4$Residuo, pch=21, bg="red", col="red")
abline(0,0, lty = 2)

```

The Durbin-Watson autocorrelation test, fifth assumption of the regression models, shows model 3 ($DWT = 2.1$) and model 4 ($DWT = 1.8$). The linear correlation between the latent variables and the dependent variable registers for the Institutional Efficiency variable model 3 ($R = 0.641$) and model 4 ($R = 0.654$) and for the Ethnic-linguistic Plurality model 3 ($R = 0.151$) and model 4 ($R = 0.007$). The variance between the factors (model 3 $chi-square = 10.41$ with 1 degree of freedom) and (model 4 $chi-square = 11.32$ with 1 degree of freedom). The results indicate absence of autocorrelation in the data, guaranteeing the integrity of the tests of significance and the confidence intervals of models 3 and 4.

```{r analyze12, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

### Test for Collinearity
vif(model_3) # variance inflation factors 
sqrt(vif(model_3)) > 3
sqrt(vif(model_3)) < 1

### Test for autocorrelation of errors 
durbinWatsonTest(model_3)

# Correlation between the dependent variable and the factor 1 - independent variable constructed
cor(analysis_data_3$`Civil Liberty 2`, analysis_data_3$`Institucional Efficiency`)

# Correlation between the dependent variable and the factor 2 - independent variable constructed
cor(analysis_data_3$`Civil Liberty 2`, analysis_data_3$`Ethnic_linguistic Plurality`)

# Test for non-constant error variance
ncvTest(model_3)

### Test for Collinearity
vif(model_4) # variance inflation factors 
sqrt(vif(model_4)) > 3
sqrt(vif(model_4)) < 1

### Test for autocorrelation of errors
durbinWatsonTest(model_4)

# Correlation between the dependent variable and the factor 1 - independent variable constructed
cor(analysis_data_4$`Civil Liberty 2`, analysis_data_4$`Institucional Efficiency`)

# Correlation between the dependent variable and the factor 2 - independent variable constructed
cor(analysis_data_4$`Civil Liberty 2`, analysis_data_4$`Ethnic_linguistic Plurality`)

# Test for non-constant error variance
ncvTest(model_4)

```

Again the sixth assumption that supports the absence of correlation between the independent variables and the error term can not be satisfied by all variables of models 3 and 4. As shown in graphs 17 and 18, in model 3 the distribution of the error term by individual parameter follows a heteroskedastic pattern for the variables institutional efficiency and accountability, but the distribution of the variables trust in government, favoritism in government decisions, and ethnic-religious plurality approximate a homoscedal pattern, already in model 4, with greater number of cases and without the variable trust in government, only the variable ethnic-linguistic plurality approaches a homoskedastic distribution, the other three variables follow a heteroscedastic pattern.

```{r analyze13, eval=FALSE, echo=FALSE}

## Non-linearity test
# component added to waste  
crPlots(model_3)

## Non-linearity test
# component added to waste 
crPlots(model_4)

```

Similar to models 1 and 2, the seventh and eighth assumptions of the regression models were satisfied from the exploratory analysis of the data, using the factorial analysis technique. We chose to use two latent variables as functional mechanisms to avoid the bias imposed by possible omitted variables and to condense information from highly correlated variables into a single factorial indicator.

Regarding the ninth assumption, which refers to the distribution of the model residues, the histograms presented in graphs 19 and 20 indicate a symmetrical distribution for the values predicted against the standardized residues in both models, which increases the validity of the inferences made by the models which use the continuous dependent variable.

```{r analyze14, echo=FALSE, message=FALSE, warning=FALSE}

par(mfrow=c(1,2))
### Evaluating waste distribution ###
# Scatter Histogram of error model 3
hist(analysis_data_3$Residuo, freq=FALSE, ylim=c(0,0.4), main="Distribution of standard error 3")
xfit <- seq(min(analysis_data_3$Residuo),max(analysis_data_3$Residuo),length=40) 
yfit <- dnorm(xfit) 
lines(xfit, yfit)

### Evaluating waste distribution ###
# Scatter Histogram of error model 4
hist(analysis_data_4$Residuo, freq=FALSE, ylim=c(0,0.4), main="Distribution of standard error 4")
xfit <- seq(min(analysis_data_4$Residuo),max(analysis_data_4$Residuo),length=40) 
yfit <- dnorm(xfit) 
lines(xfit, yfit)


```

Finally, in model 4, we look at the tenth assumption regarding the Central Limit Theorem that proposes to increase the number of cases against the estimated parameters, in annex the table 11 indicates that when we increase the number of cases from 51 in model 2, to 127 in model 4, removing the variable trust in government, the result indicates that the explanatory power of model 4 is greater when compared to the previous one in terms of $R-adjusted$, it is observed that in this model the favoritism variable in the government decisions has effect significant difference on civil liberty, unlike model 2 where there is no significance for this variable, in addition, the standard error of the statistically significant independent variables was reduced, therefore, the increase in the number of cases had a great influence on the robustness of the results of model 4.



---

References {#references .unnumbered}
==========

Annex 1
=============

```{r analyze15, eval=FALSE}

\begin{center} Tabel 1 - Indicators \end{center}

|          **Factor 1**          |          **Factor 2)**                   
|:-------------------------------|---------------------------------------:| 
| ** Institutional Efficiency    | Ethnic-linguistic Plurality
|:-------------------------------|---------------------------------------:| 
| X1 (Corruption Control)        | X6 (Cultural Diversity)
| X2 (Ethic Firms Behavior)      | X7 (Ethnic Fracionalization)
| X3 (Property Rigths)           | X8 (Linguistic Fracionalization)
| X4 (Political Stability)       |
| X5 (Functioning of government) |
|:-----------------------------------------------------------------------:|
*Source: By the author (2019)*

```


```{r analyze16, eval=FALSE}

\begin{center} Tabel 2 - Descriptive Statistics \end{center}

|         **Variables**           | **Mean**  | STD                   
| ------------------------------- |:---------:|-----------:| 
| X1 (Corruption Control)         | 0.08      | 0.96
| X2 (Ethic Firms Behavior)       | 4.11      | 0.96
| X3 (Property Rigths)            | 4.26      | 1.01
| X4 (Political Stability)        | -0.06     | 0.98
| X5 (Functioning of government)  | 4.99      | 2.51
| X6 (Cultural Diversity)         | 0.31      | 0.20
| X7 (Ethnic Fracionalization)    | 0.47      | 0.26
| X8 (Linguistic Fracionalization)| 0.39      | 0.28
|:-------------------------------------------------------:|
*Source: By the author (2019)*

```


```{r analyze 17, eval=FALSE}

\begin{center} Tabel 3 - Sample Suitability Tests \end{center}

| **Test** | **Value observed** |                   
| ---------|:------------------:| 
| **KMO**  | 0.860
| **BTS**  | 36
| **SIG**  | 0.01
|:----------------------------: | 
*Source: By the author (2019)*

```

```{r analyze 18, eval=FALSE}

\begin{center} Tabel 4 - Communalities \end{center}

|                 | **Components**           
| -------------------------------------------
|         **Variable**            |  Inicial | **Extracted** |
|:----------------------------------------------------------:| 
| X1 (Corruption Control)         | 1.000    |     0.911     |
| X2 (Ethic Firms Behavior)       | 1.000    |     0.890     |
| X3 (Property Rigths)            | 1.000    |     0.845     |
| X4 (Political Stability)        | 1.000    |     0.603     |
| X5 (Functioning of government)  | 1.000    |     0.665     |
| X6 (Cultural Diversity)         | 1.000    |     0.843     |
| X7 (Ethnic Fracionalization)    | 1.000    |     0.829     |
| X8 (Linguistic Fracionalization)| 1.000    |     0.772     |
|:----------------------------------------------------------:|
*Source: By the author (2019)*
```


```{r analyze 19, eval=FALSE}

\begin{center} Tabel 5 - Cumulative Variance \end{center}

|                                 |    **Accumulated Values**        |    Square load rotating sums     |
| ------------------------------------------------------------------------------
|          **Component**          | StD  | Variance % | Acumulated % | STD  | Variance % | Acumulated % |
|:-----------------------------------------------------------------------------------------------------:| 
| X1 (Corruption Control)         | 2.06 | 53         | 53           | 4.64 | 51         | 51           |
| X2 (Ethic Firms Behavior)       | 1.45 | 26         | 79           | 2.57 | 28         | 80           |
| X3 (Property Rigths)            | 0.69 | 0.6        | 85           | -    | -          | -            |
| X4 (Political Stability)        | 0.64 | 0.05       | 90           | -    | -          | -            |
| X5 (Functioning of government)  | 0.60 | 0.04       | 95           | -    | -          | -            |
| X6 (Cultural Diversity)         | 0.41 | 0.02       | 97           | -    | -          | -            |
| X7 (Ethnic Fracionalization)    | 0.35 | 0.01       | 99           | -    | -          | -            |
| X8 (Linguistic Fracionalization)| 0.22 | 0.01       | 100          | -    | -          | -            |
|:-----------------------------------------------------------------------------------------------------:|
*Source: By the author (2019)*
```   

```{r analyze20, eval=FALSE}

\begin{center} Tabel 6 - Rotated Matrix Component (varimax) \end{center}

|    Chi-squared: 97.63  DF: 26    |  **Components**   | Uniquiness        |
| -------------------------------------------------------------------------|
|     **Independent Variables**    |   1     |    2    |                   | 
|:------------------------------------------------------------------------:| 
| X1 (Corruption Control)          | 0.982   | -0.160  | 0.010             | 
| X2 (Ethic Firms Behavior)        | 0.911   | -       | 0.165             | 
| X3 (Property Rigths)             | 0.853   | -       | 0.302             | 
| X4 (Political Stability)         | 0.800   | -0.164  | 0.333             | 
| X5 (Functioning of government)   | 0.758   | -0.164  | 0.339             | 
| X6 (Cultural Diversity)          | 0.751   | 0.987   | 0.025             | 
| X7 (Ethnic Fracionalization)     | -0.203  | 0.760   | 0.395             | 
| X8 (Linguistic Fracionalization) | -0.203  | 0.707   | 0.497             |
|:------------------------------------------------------------------------:|
*Source: By the author (2019)*
  
```

```{r analyze21, eval=FALSE}

\begin{center} Tabel 7 - Mean variation of factors \end{center}

|             **N: 127**           | **min** |  **max** |   
|:-----------------------------------------------------:| 
| F1 (Institutional Efficiency)    | -3.70   | 4.37     |
| F2 (Ethnic-linguistic Plurality) | -2.87   | 3.32     | 
|:-----------------------------------------------------:|
*Source: By the author (2019)*

```

Annex 2
================

Dependent variable:
Y) Civil liberty 1: (fh_cl) categorical dependent variable that measures rights such as freedom of expression and belief, association and organization, force of law, and autonomous individualism without state interference on the particular. The indicator ranges from 1 (very free) to 7 (less free). Models 1 and 3, respectively, contain information for 132 and 51 countries.

Y) Civil liberty 2: (eiu_cl) continuous dependent variable that measures civil liberties from freedom of expression and press; freedom of religion; freedom of assembly and association; and the right to due process. The indicator ranges from 1 to 10, where the freest country receives the maximum score = 10, and the least free the minimum score = 1. Models 2 and 4, respectively, contain information for 132 and 51 countries.
Independent variables:

i) X1 = trust in the government, it has information for 64 countries and it measures the average trust in the government, the indicator varies from 1 (without trust) to 4 (total trust), data were collected between 2009 and 2014;

ii) X2 = favoritism in official government decisions, has information for 148 countries, measures the extent of government power, the power of government officials to favor companies and individuals in agreements, for example, acting as a circle of influence in making decisions on policies and contracts. The indicator ranges from 1 (always show favoritism) to 7 (never show favoritism), data were collected from 2011 to 2013;

iii) X3 = transparency accountability, has an index composed of 16 indicators that measure transparency in government actions (six of them measure the free undertaking, four for fiscal transparency, and six for political constraints). The base year is 1980, the transparency index had information for 115 countries in 1980, this figure rose to 189 countries by the end of 2010.

iv) X4 = Factor 1 - institutional efficiency

v) X5 = Factor 2 - ethnic-linguistic plurality
(QOG, Codebok, 2016).



```{r analyze22, eval=FALSE}

\begin{center} Tabel 8 - Regression Model Variables \end{center}

| **Dependent Variable: Civil Liberty 1 - 2**   
| ---------------------------------------------------|
| **Independent Variables**  
|:--------------------------------------------------:| 
| X1 (Confidence in Government)                      |
| X2 (Favoritism in Government officials decisions)  |
| X3 (Accountability)                                |
| X4 (Institutional Efficiency)                      |
| X5 (Ethnic-linguistic Plurality)                   |
|:-------------------------------------------------: |
*Source: By the author (2019)*
```  

```{r analyze23, eval=FALSE}

\begin{center} Tabel 9 - Descriptive Statistics \end{center}
  
|                   **VariableS**                   | **Mean** |  min  |  máx  |                 
|:-------------------- |------------------------------------------------------:| 
| X1 (Confidence in Government)                     | 2.25     | 1.68  | 3.32  |
| X2 (Favoritism in Government officials decisions) | 2.99     | 1.85  | 5.41  |
| X3 (Accountability)                               | 56       | 11    | 80    |
| X4 (Institutional Efficiency)                     | -0.40    | -3.70 | 4.37  |
| X5 (Ethnic-linguistic Plurality)                  | 0.07     | -2.87 | 3.32  |
|:----------------------------------------------------------------------------:|
*Source: By the author (2019)*

```


```{r analyze24, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Viewing models 1 to 4
stargazer(model_1, model_2, model_3, model_4, 
          type = "text", title = "Model Results", style = "ajps", align=TRUE, no.space=TRUE, p.auto=FALSE)

```




